---
title: "R Assignment 1 EDA"
author: "name"
date: "`r Sys.Date()`"
output: pdf_document
execute: 
  echo: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(openintro)
```

## Assignment: 

You will do an exploratory data analysis on the yrbss survey data using R. (See the template in the MTH_142_files you uploaded to Rstudio). Click here for some documentation.


You will turn in a pdf output of your work.

Introduce the data with a summary. (50 words)
Calculate at least two summary statistics (center, spread, count, or a table).
Create a visualization from the data.
Write a brief explanation of what you learned from your statistics and visualization. (50 - 100 words)

## Introduce your data


## Calculate two Summary Statistics

```{r statistic_1}
# The code for your 1st statistic goes here

```

```{r statistic_2}
# The code for your 2nd statistic goes here

```

## Make a plot

```{r}
# Write some ggplot() code for a plot here. 

```

## Conclusion

