---
title: " R Assignment 6 "
subtitle: "Test for two means" 
author: "optional name"
date: "`r Sys.Date()`"
output: pdf_document
execute: 
  echo: true
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(eval = TRUE)
library(tidyverse)
library(openintro)

```

# Intro/Goal

In this assignment you will carry out a hypothesis test for the difference between two means. Specifically you will see if the average birthweight of a baby born from a smoking mother is different than the birthweight of a baby born from a non-smoking mother. 

As with R Assignment 4 you will do an EDA, check the conditions and finally come to a conclusion. You will also have to consider which testing error is possible. Along the way I will introduce you to a new function called t.test() which will make computing the hypothesis test easier.

The data is from births14. There are many variables in this data, but you will be studying two: weight and habit. In the code chunk below I've removed all other variables. 

```{r}
births14 = births14 |> select(weight,habit)


```

## Do an exploritory analysis. 

At minimum you should make a boxplot to check for extreme outliers. You should plot weight, but fill by habit. (see R Assignment 1 or my example video on how to do this) You might want to check for other information that might be necessary for conditions.  

```{r}

```

## Check conditions

Check the conditions of the test. You can find these conditions in chapter 20.



## Write out the hypothesis notation. Include a significance level.

Edit the values below to match this test. I've done the null hypothesis for you. You need to edit the alternative hypothesis and alpha. 

$$
H_o: \mu_{smoker} = \mu_{non\_ smoker} \\ H_a: \mu \ne 0 \\ \alpha = 1

$$
## Do the test and state your p-value.
   
If you'd like to do a difference of means test like you did in the homework that is fine and you can disregard the t.test() function below.    
    
Otherwise use t.test() to have R calculate the p-value for you. State your p-value.
note: t.test() needs four arguments for this test. 

-   formula of the form = numeric variable ~ categorical variable 
-   data = births14
-   alternative = the alternative hypothesis (use "less", "greater","two.sided"). 
-   correct = FALSE (I've done this for you)
  
```{r}
# This line of code turns off scientific notation, you don't need to do anything with it. 
options(scipen = 100)

t.test(formula =  , data =  , alternative = "", correct = FALSE)

```

## Conclusion

Write a conclusion, be sure to state your pvalue and state what it means in terms of the weights of babies from smoking and non smoking mothers. Make sure to discuss whether the conditions were met and any issues that might cause your conclusion to be incorrect. 


## Discuss which testing error (I or II) may be possible.

Note: this does not mean your test is invalid. We are dealing with probabilities so it is always possible for for our conclusion to be incorrect. 

