---
title: "MTH 142: R Assignment 3"
author: "optional name"
date: "`r Sys.Date()`"
output: pdf_document
execute: 
  echo: true
---

```{r setup, include=FALSE}
library(openintro)
library(tidyverse)

# This is the bring up the documentation for the smallpox data. 
?smallpox

#This will put the smallpox_no_vaccine data into your environment
smallpox_no_vaccine <- smallpox |> filter(inoculated == "no") |> select(result)
```

It is claimed that smallpox has a 30% mortality rate for those that are not vaccinated against it. In this 3rd R assignment you will test a sample of people who were not vaccinated but were exposed to smallpox to see if that claim is true. The data comes from the smallpox data frame and you can read about it in the help menu after running the chunk above. 

However you will not use the smallpox dataframe as some people within that data are vaccinated. In the chunk above I filtered the data to only include unvaccinated people. For this assignment use the small_pox_no_vaccine data frame. smallpox_no_vaccine has one column called result and it shows whether the person lived or died.  

## An exploratory data analysis.
    
You should have a look at the data, perhaps make a bar graph or table to see the current distribution of the data. You may want to reference the lecture on categorical variables from week 1, look at a cheat sheet, review R_Assignment_1, or ask questions in the forum on how to do this. 

```{r}


```

    
## Check the conditions.*

Every test has conditions. You need to check that the conditions for this test are met by examining the data. This is a one proportion test so you should consult chapter 16 of IMS. You will likely need to produce code that counts the levels (lived or died) of the variable result. I do this with `count()` in the video. Be sure to type out whether the conditions are met or not and justify. Continue with the test even if conditions are not met. 

```{r}
count(   ,   )
```


## Write out the hypothesis notation. Include a significance level.

In this part you will edit the numbers below to display the mathematical symbols for the hypothesis test. Verify your notation by looking at the preview. Note: the symbols are all correct, you only need to change the values (Change the 0 and 1s).Leave the dollar signs alone.  

$$H_o: p = 1 \\ H_a:p \ne 0 \\ \alpha = 1$$

## Do the test and state your p-value.
    
When doing your homework you found the theoretical distribution and calculated probabilities with pnorm(), there is a function in R that will do all this for you. It is called prop.test(). I describe it below. If you'd like you can go back to using pnorm() instead. 
    
Use prop.test() to have R calculate the p-value for you. State your p-value.
note: prop.test() needs five arguments. 
-   x = number of successes, people who favor a quarantine. 
-   n = total sample size
-   p = the assumed population parameter (This should match your hypothesis notation above)
-   alternative = the alternative hypothesis , could be "less than" or "greater than" (I've set this for you). 
-   correct = FALSE 
  
```{r}
# This line of code turns off scientific notation, you don't need to do anything with it. 
options(scipen = 100)

# This is an outline for the function prop-test, fill in the blanks. Post questions to the forums. 
prop.test(x =  , n = , p = , alternative = "two.sided", correct = FALSE)

```

## Write the Conclusion.

Considering your p-value and alpha. State whether is it likely that true population parameter is 0.50 based on the data. Be sure to state both the pvalue and alpha. 


## Discuss which testing error (I or II) may be possible.

Read chapter 14 or ask in the forums for notes on decision errors. The talk about which you may have made. 

Note: this does not mean your test is invalid. We are dealing with probabilities so it is always possible for for our conclusion to be incorrect. 
