---
title: "Continuous RV"
author: "Schwab"
format: 
  revealjs:
    theme: beige
editor: visual
---

## Reading

[Read chapter 3.5 from Openintro statistics.](/2019.openintro.statistics.pdf)

## Continuous RV

::: incremental
-   A continuous random variable can take any numeric value.

-   (Discrete can only take integers.)

-   X- Face value of a die

-   Y- Sum of two die.

-   X and Y are **discrete** r.v. because their outcomes are whole numbers.
:::

## Height

X- height (inches)

X is a continuous random variable.

(people usually report is as a discrete RV.)

## Tangent: Precision matters

Who is 65 inches tall?

None of us really know how tall we are exactly.

## Convenience Sample on heights

Last semester my in person class told me how tall they are. I made a histogram of their heights.

```{r}
library(tidyverse)

heights <- c(65L, 62L, 69L, 68L, 63L, 61L, 61L, 64L, 63L, 62L, 71L, 66L, 63L, 68L, 67L, 70L, 65L, 61L, 64L, 69L, 67L, 67L, 70L, 64L, 61L, 64L, 64L, 69L, 69L, 59L, 63L, 66L, 64L, 62L)

heights <- data_frame(heights)


ggplot(data = heights)+
  geom_histogram(mapping = aes(heights), bins=12, color="orange", fill= "gray40")+
  theme_classic()+
  xlab("Heights Inches")+
  ylab("Number of Students")+
  labs(title = "Heights from students Fall 2023")
```

## Probabilities from our sample

There are 34 students in my sample (n = 34).

Treat the data as discrete.

::: incremental
Estimate the following probabilities.

-   P( X \< 65 )
-   P( 60 \< X \< 63 )
-   P( X = 65 )
:::

```{r}
table(heights)
```

## Cont. Height Density Plot

We get slightly different answers if we think about the data as continuous.

Let X  - A Student's Height

```{r}

ggplot(data = heights)+
  geom_density(mapping = aes(heights), fill= "gray40", alpha = 0.4)+
  theme_classic()+
  xlab("Heights Inches")+
  ylab("Number of Students")+
  labs(title = "Heights from students Fall 2023")

```

## Probability Height Questions

Let's pretend our sample is representative of all students

::: incremental
-   P( X \< 65 )
-   P( 60 \< X \< 63 )
-   P( X = 65 )

Cannot be equal to exactly 1 number with continuous distributions. There is an $\infty$ number of values.
:::

## You try.

[3.37](https://www.biostat.jhsph.edu/~iruczins/teaching/books/2019.openintro.statistics.pdf#chapter.3)
