---
title: "Continuous RV"
author: "Schwab"
format: 
  revealjs:
    theme: beige
editor: visual
---

## Continuous RV

::: incremental
-   Remember continuous means we can take any numeric value.

-   (Discrete means we can take integers.)

-   X- Face value of a die

-   Y- Sum of two die.

-   X and Y are **discrete** because they deal while whole numbers
:::

## Height

X- height (inches)

X is a continuous random variable.

(people usually report is as a discrete RV.)

## Tanget: Precision matters

Who is 65 inches tall?

None of us really know how tall we are exactly.

## Convience Sample on heights

Remember when you all told me your heights? I made them into a histogram.

```{r}
#| echo: FALSE

library(tidyverse)

heights <- c(65L, 62L, 69L, 68L, 63L, 61L, 61L, 64L, 63L, 62L, 71L, 66L, 63L, 68L, 67L, 70L, 65L, 61L, 64L, 69L, 67L, 67L, 70L, 64L, 61L, 64L, 64L, 69L, 69L, 59L, 63L, 66L, 64L, 62L)

heights <- data_frame(heights)


ggplot(data = heights)+
  geom_histogram(aes(heights), bins=12,color="orange", fill= "gray40")+
  theme_classic()+
  xlab("Heights Inches")+
  ylab("Number of Students")+
  labs(title = "Heights from SDS 220 Fall 2023")
```

## Probabilities from our sample

Hint: n = 34 students responded to the survey.

::: incremental
-   P( X \< 65 )
-   P( 60 \< X \< 63 )
-   P( X = 65 )

Cannot be equal to exactly 1 number with continuous distributions. $\inf$ number of values.
:::

## Cont. Height Histogram

```{r}

ggplot(data = heights)+
  geom_histogram(aes(heights), stat = "density", bins=10,color="orange", fill= "gray40")+
  theme_classic()+
  xlab("Heights Inches")+
  ylab("Number of Students")+
  labs(title = "Heights from SDS 220 Fall 2023")

```

## Probability Height Questions

Let's pretend our sample is representative of all Smithies

::: incremental
-   P( X \< 65 )
-   P( 60 \< X \< 63 )
-   P( X = 65 )

Cannot be equal to exactly 1 number with continuous distributions. $\inf$ number of values.
:::

# Just one problem

[3.37](https://www.biostat.jhsph.edu/~iruczins/teaching/books/2019.openintro.statistics.pdf#chapter.3)
