---
title: "Hypothesis Testing"
author: "Schwab"
format: 
  revealjs:
    theme: beige
editor: visual
---

## Example: Ebola

::: incremental
-   Pretend: You live in NYC and you all are perfectly representative of denizens there.

-   Do you favor a "mandatory 21-day quarantine for anyone who has come in contact with an Ebola patient?"

-   Class convenience sample. This will give us $\hat{p}$.
:::

[Ebola in New York City](https://en.wikipedia.org/wiki/Ebola_virus_cases_in_the_United_States)

## Survey Data

Our question: We want to know if New Yorkers have changed their stance on quarantine of Ebola after the COVID-19 pandemic.

## The 2014 data.

We'll use the `ebola_survey` data as the population to test against. This will be our $p$.

```{r}
library(tidyverse)
library(openintro)
count(ebola_survey,quarantine)
```

This is a reasonable estimate for *p* because...

## p vs $\hat{p}$

::: incremental
-   $\hat{p}$ - the sample statistic. It is a proportion in this case.

    -   This is how this subset of New Yorkers feel about quarantine.

-   p - The assumed population parameter, also a proportion.

    -   This is how all New Yorkers feel about a quarantine.
:::

## The question.

\> If the sentiment from 2014 New Yorkers is unchanged to today, what is the probability we would have gotten the $\hat{p}$ we found?

::: incremental
-   Is it likely or unlikely

    -   Likely will be anything more than $\alpha = 0.05$

    -   If unlikely we reject $H_0$
:::

## Notation

Hypothesis testing with Proportions

$$
H_o: p = 0.82 \\
H_1: p \ne 0.82
$$

Null and Alternative hypotheses.

We assume the null hypothesis is true and build a theoretical sampling distribution from that.

## The Central Limit Theorem

This is how we build the sampling distribution.

> If we look at a proportion (or difference in proportions) and the scenario satisfies certain conditions, then the sample proportion (or difference in proportions) will appear to follow a bell-shaped curve called the *normal distribution*.

## Condition on $H_0$

Proportions:

-   Success Failure condition: $n(p_0) > 10$ and $n(1-p_0) > 10$

-   Large Independent Samples (n\>30)

## Draw the Sampling Distribution

$$ p \sim N(p_0,SE)$$

With $$SE= \sqrt{\frac{p_0(1-p_0)}{n}}$$

## Conclusion

What is the probability of getting the $\hat{p}$ from our sample?

```{r}
#| eval: false
#| echo: true

pnorm(q = p-hat, mean = p, sd = SE, lower.tail = )
```

# Board work

## Ex 1. Exclusive relationships

What is the proportion of college students that have had more than 1 `exclusive_relationships`?

```{r}
exclusive_relationship |>
  count(num > 1)

```

-   Check Conditions.

-   Do the test.

## Problems

[Chapter 16: 13, 14, 15](https://openintro-ims.netlify.app/inference-one-prop#chp16-exercises)

## iPhones

In a sample of 300 students, 68% said they own an iPod and a smart phone. Compute a 97% confidence interval for the true percent of students who own an iPod and a smartphone.[^1]

[^1]: This and the previous problem were from a text called [Introductory Statistics](https://openstax.org/books/introductory-statistics/pages/8-3-a-population-proportion)

<!-- ## z-scores you should know: -->

<!-- | CI    | Z-score          | -->

<!-- |-------|------------------| -->

<!-- | 90%   | $z_{0.05}=1.645$ | -->

<!-- | 95%   | $z_{0.025}=1.96$ | -->

<!-- | 99.7% | $z_{0.005}=$     | -->

<!-- : z-scores you should know. -->

<!-- ## You should know: -->

<!-- How to calculate and interpret the theoretical CI for a single proportion and mean. -->

<!-- The difference between the data, sample and bootstrap distributions. -->

<!-- How to make a bootstrap distribution and use it to find a CI. -->

<!-- ## Hypothesis Testing -->

<!-- This is the last topic on exam 2. -->

<!-- Read chapters 16 and 19 before Monday and Wednesday. -->

<!-- ## Hypothesis Testing -->

<!-- In this talk I'll introduce some hypothesis testing for a single mean and proportion. -->

<!-- We'll also do some examples. -->

<!-- # A single Mean -->

<!-- ## Example 1 -->

<!-- Pretend is known that the average time it take a runner to finish a 5k is 35 minutes. -->

<!-- This is the made up average for all 5ks for which there is data. -->

<!-- We can write: -->

<!-- $H_0: \mu = 35$ -->

<!-- Its the status quo. -->

<!-- ## Alternative Hypothesis -->

<!-- I've run the `get_it_dunn_5k` and it is challenging. -->

<!-- I think the average time will be different. -->

<!-- We can write that as: -->

<!-- $H_a: \mu \ne 35$ -->

## Test the hypothesis

Plan:

-   State Hypothesis, set $\alpha$

-   Collect Data and Check conditions

-   Find Assume the null hypothesis is true

-   Find the p value that goes with $\hat{p}$

-   Conclusion

<!-- ## Conditions -->

<!-- In general when using the Central Limit Theorem to make CIs based on the Theoretical Sampling Distribution we want a few things to be true: -->

<!-- For a mean: -->

<!-- -   Large Independent Samples (n\>30) -->

<!-- -   No particularly extreme outliers. -->

<!-- ## Large Independent Sample? -->

<!-- ```{r} -->

<!-- library(openintro) -->

<!-- library(tidyverse) -->

<!-- get_it_dunn_run |> -->

<!--   filter(race == "5k") |> -->

<!--   group_by(race) |> -->

<!--   summarise( n=n()) -->

<!-- ``` -->

<!-- It is large and extremely likely it is independent, so let's assume that. -->

<!-- ## Check outliers and normality? -->

<!-- ```{r} -->

<!-- get_it_dunn_run |> -->

<!--   filter(race == "5k") |> -->

<!--   ggplot(aes(x=run_time_minutes))+ -->

<!--     geom_boxplot() -->

<!-- ``` -->

<!-- One not extreme outlier. This looks ok enough. -->

<!-- ## Make a 95% CI. -->

<!-- Use $z_{0.025} = 1.96$ -->

<!-- ```{r} -->

<!-- library(openintro) -->

<!-- library(tidyverse) -->

<!-- get_it_dunn_run |> -->

<!--   filter(race == "5k") |> -->

<!--   group_by(race) |> -->

<!--   summarise(mean = mean(run_time_minutes), sd= sd(run_time_minutes), n=n()) -->

<!-- ``` -->

<!-- ## Things to note: -->

<!-- Note: we have not proved anything. Recall our three reasons why -->

<!-- $H_0: \mu$ and ð‘¥ could be dissimilar: -->

<!-- -   Sample is not representative. -->

<!-- -   Random chance. -->

<!-- -   H_0 is false. -->

<!-- The first option is unlikely because we are responsible statisticians. We -->

<!-- have found that random chance (second option) is very unlikely (our -->

<!-- confidence level is usually high). So it seems that the last option is the -->

<!-- most plausible. -->

<!-- ## Rejection Region -->

<!-- How big is our rejection region? -->

<!-- This is called our significance level. -->

<!-- It is given the notation $\alpha = 0.05$ -->

<!-- ## Otherways to do hypothesis testing. -->

<!-- -   Confidence Intervals -->

<!-- -   p-values (coming next week) -->

<!-- -   critical values -->
