---
title: "Confidence Intervals with a proportion"
author: "Schwab"
format: 
  revealjs:
    theme: beige
editor: visual
---

```{r}
#| echo: false
#| include: false


library(tidyverse)
library(openintro)
```

## Practice Making CI with proportions

## Find the confidence interval for:

1.  The true number of US women who were married before age 30.

```{r}
#| code-fold: true

library(openintro)

age_at_mar <- age_at_mar |> 
  mutate(over_30 =
           case_when(
             age > 30 ~ "Over_30",
             TRUE ~ "Under_30"
           ))
head(age_at_mar, 2)

```

2.  The true proportion of women who smoke while pregnant in the US.

```{r}
head(births14)

```

## CI mathematically

We build the CI around the sample proportion.

Assume

$$
\hat{p} \sim N(p,se)
$$

1.  We need $\hat{p}$ and $\text{se}= \sqrt{\frac{p(1-p)}{n}}$

2.  We also need the z score that goes with the confidence interval we are looking for.

3.  $\hat{p} \pm z_{score} \times se$

## Remember these z-scores

| CI    | Z-score           |
|-------|-------------------|
| 90%   | $z_{0.05}=1.645$  |
| 95%   | $z_{0.025}=1.96$  |
| 99.7% | $z_{0.005}=2.968$ |

: z-scores you should know

## How to find zscore with `qnorm()`

`qnorm()` finds the quantile that goes with a pvalue (probability)

```{r}
qnorm(p = 0.975, mean = 0, sd = 1 )
```

`pnorm()` finds a pvalue from a quantile.

```{r}
pnorm(q=1.959964 , mean = 0, sd = 1 )
```

## The distribution of sampling proportions

Compare vs. bootstrap distributions.

## Let's make a 95% CI

Find these: $\hat{p}$ and $\text{se}= \sqrt{\frac{p(1-p)}{n}}$

$\hat{p} \pm z_{score} \times se$

## Assumptions for a single parameter

Independent and "large" sample.

Success-Failure Condition

Why do we need to assume this:

$$
\hat{p} \sim N(p,se)
$$

## Board Practice

We found these with simulation, find them now with a mathematical model. Use a 88% CI.

1.  The true number of US women who were married before age 30.
2.  The true proportion of women who smoke while pregnant in the US.

# The Central Limit Theroem

# Notation

$$ \hat{p} \sim N(p,SE)$$

With $SE= \sqrt{\frac{p(1-p)}{n}}$

or

$SE= \sqrt{\frac{\hat{p}(1-\hat{p})}{n}}$

# Rules of thumb

::: {incremental}
-   The operations need to be independent.

-   n should be large

-   Success-Failure Condition

    -   $np >=10$ and

    -   $n(1-p) >= 10$
:::

## Quarantine

We ask 1042 New Yorkers if they are for quarantining peope who have been exposed to ebola. The results are below.

```{r}
count(ebola_survey,quarantine)
```

-   Write the notation for the theoretical sampling distribution.

-   Make a 90, 95 and 99.7% CI for those against and interpret.

## Exclusive relationships

What is the proportion of college students that have had more than 1 `exclusive_relationships`?

```{r}
exclusive_relationship |>
  count(num > 1)

```

-   Check Conditions.

-   Write the notation for the theoretical sampling distribution.

-   Make a 90, 95 and 99.7% CI for those against and interpret.

## iphones

b.  In a sample of 300 students, 68% said they own an iPod and a smart phone. Compute a 97% confidence interval for the true percent of students who own an iPod and a smartphone.[^1]

[^1]: This and the previous problem were from a text called [Introductory Statistics](https://openstax.org/books/introductory-statistics/pages/8-3-a-population-proportion)

## Worse off?

Every week the Yougov/The Economist do a survey from a representative sample of about 1500 adults. [methodology](https://today.yougov.com/about/panel-methodology) On March 4, 2024 people were asked about their personal finances. See the results [here](https://d3nkl3psvxxpe9.cloudfront.net/documents/econTabReport_TYCEeIV.pdf#table.143).

Find a CI for the true proportion of Americans who feel they are worse off financially than a year ago.

```{r}
number_worse_off <- round(0.42*1556)
not_worse_off <- 1556 -number_worse_off

# This makes a data frame that mirrors the number of people who were worse off or not from the poll. 
poll <- tibble(
  worse_off = c(rep("Yes", number_worse_off), rep("No", not_worse_off))
)
```

## Problems from the text

Do these with simulation and then with math. [Chapter 6: 20, 21, 22](https://today.yougov.com/about/panel-methodology)
