---
title: "ANOVA part 1"
author: "Schwab"
format: 
  revealjs:
    theme: beige
editor: visual
execute: 
  echo: true
---

## Reading

[Read Chapter 22 before hand.](https://openintro-ims.netlify.app/inference-many-means)

## Libraries

```{r}
library(openintro)
library(tidyverse)
```

## Comparing multiple groups

We use **An**alysis **o**f **Va**riance for testing if there is a difference between many means.

## New distribution: F

The F distribution looks like:

```{r curve-F}
#| echo: False

curve(
  df(x, df1=1, df2=2), 
  from=0, to=5)
```

## p-value and CI

We find the p-value in the right tail.

There is no real confidence interval.

## Conditions:

-   **the observations are independent within and between groups,**

<br>

-   the responses within each group are nearly normal, and\*

<br>

-   the variability across the groups is *about* equal.

## Rule of Thumb

<br>

"The variability across groups is equal" condition has a rule of thumb.

If the largest SD/ smallest SD is between 0.5 and 2.

$$
0.5\le \frac{\sigma_{max}}{\sigma_{min}}\le 2
$$

## A test

There are three classes that took a midterm. We want to know if the exam average for any of the classes is different.

```{r}
ggplot(data=classdata) +
  geom_boxplot(
    aes(x= m1, color = lecture)    )
```

## Summary stats

Is the variance about the same?

```{r}

group_by(.data = classdata, lecture) |>
  summarise(mean = mean(m1), sd = sd(m1), count = n())
```

## Notation:

<br>

$H_0: \mu_1 =\mu_2 =\mu_3$

$H_a:$ At least one mean is different.

$\alpha = 0.05$

## The test computation

Is done by R with `aov()` . We will not be doing this test by hand.

```{r}
results <- aov( m1 ~ lecture, data = classdata )
summary(results)
```

## Conclusion:

Reject the null, one of the exam averages is different than the other two.

## A flower's sepal

![](https://upload.wikimedia.org/wikipedia/commons/7/7f/Mature_flower_diagram.svg)

## Try `Sepal.Length` from iris.

Is it appropriate to do ANOVA on `Sepal.Length`?

```{r}
#| echo: False

group_by(.data= iris, Species) |>
   summarise(mean = mean(Sepal.Length), sd = sd(Sepal.Length))
```

```{r}
#| echo: False

iris |>
  ggplot(aes(Sepal.Length, color = Species))+
  geom_boxplot()

```

## Look at the graph

Is there one mean that's different?

```{r}

iris |>
  ggplot(aes(Sepal.Length, color = Species))+
  geom_boxplot()
```

## Do the test

-   Notation

-   Test

-   Conclusion

```{r}
results <- aov( Sepal.Length ~ Species, data = iris )
summary(results)
```

## Problems 5 or 9

[Link to exercises](https://openintro-ims.netlify.app/inference-many-means#chp22-exercises)

## Show how to do HW problem 3 and 4
