---
title: "Difference of Means"
author: "Schwab"
format: 
  revealjs:
    theme: beige
editor: visual
execute: 
  echo: true
---

## Warm up

Find a partner (or two) and grab some chalkboard

Do this hypothesis test.

The average height of Smith students is thought to be 64 inches. Our class has an average height of 65 inches with a standard deviation of 3.2. There were 34 students in our sample.

Test the hypothesis that the true mean is 64 twice

1.  once using the observed statistic to find a p-value.

2.  once with the test statistic to find the same p-value.

## warm up again with R

```{r}
library(tidyverse)
library(googlesheets4)

# Making CI from some sample proportions.

class_heights <- read_sheet("https://docs.google.com/spreadsheets/d/1x9srZXDoYzKGew1f-y4brMbN8dl4QM66pc-dmCBWHQY/edit?usp=sharing",
                            sheet = "220-01",
                            range = "1:35")|>
  select(height)

```

```{r}
t.test(class_heights$height, mu=64)
```

## Side note: `prop.test()`

There is also a `prop.test()` function that will test a proportion hypothesis from data based on mathematical formula (as opposed to bootstrapping).

## New test: Difference of two means

Another hypothesis test.

-   Conditions

-   Distributions

-   New Standard Error

-   New df

# Conditions

Independence Extended

Normality (no extreme outliers)

# Compute the df

**with no computer**

t - distribution with the smaller of $n_1-1$ and $n_2-1$ degrees of freedom.

<br>

**with R**

t - distribution, let R find the degrees of freedom with `t.test()`.

# SE

$$
SE = \sqrt{\frac{\sigma^2_1}{n_1}+\frac{\sigma^2_2}{n_2}}\\
\text{or}\\
SE = \sqrt{\frac{s^2_1}{n_1}+\frac{s^2_2}{n_2}}
$$ We rarely know $\sigma$ so the second standard error is more often used.

# Test Statistic

$$
\text{Test Statisitc} = \frac{\bar{x}_1 - \bar{x}_2 -0}{SE }
$$

(It looks just like the difference of proportions, but with means)

# Let's do number 18 together.

[Chapter 20](https://openintro-ims.netlify.app/inference-two-means#chp20-exercises)

1.  CI for the difference of means
2.  Hypothesis Test

# Grab a partner

20.17 on the board.

# Together t.test()

Is there a difference in the absenteeism of students from New South Wales based on reported gender?

-   `absenteeism`

-   check conditions make boxplot

-   write hypothesis

-   use `t.test( y\~x, alternative= "t", data=)`

# A second test

Is there a difference in the absenteeism of students from New South Wales based on ethnicity?
