---
title: "Exploring Categorical Data"
subtitle: "IMS, Ch. 4"
author: "Nic Schwab"
format:
  revealjs: 
    slide-number: true
    theme: beige
execute: 
  echo: true
---

## Read before Lecture

[Chapter 4 Exploring Categorical Data](https://openintro-ims.netlify.app/explore-categorical)

## Confounding Variables

For each of the following pairs of variables, a statistically significant positive relationship has been observed. Identify confounding that might cause the [spurious correlation](https://en.wikipedia.org/wiki/Spurious_relationship).

::: incremental
-   The amount of ice cream sold in New England and the number of deaths by drowning
-   The number of doctors in a region and the number of crimes committed in that region

:::

## Explore Categorical Data

We'll be exploring categorical data using R

## Chapter 4

Exploring Categorical Data:

```{r}
library(tidyverse)
library(openintro)

# You should read and view the documentation of the data as a first step to exploring it. 
?assortive_mating
assortive_mating
```

## Exploring categorical data {#sec-explore-categorical}

Let's explore **categorical** data using summary statistics and visualizations.

## Tables

Tables are also helpful in understanding categorical data

```{r}
table(data = assortive_mating)
```

## Prop tables

```{r}
# Here I am saving the table as the variable my_table
# I am telling table I specifically want to look at the male and female variables.

my_table <- table(assortive_mating$self_male, assortive_mating$partner_female)

prop.table(my_table)
```

## Margins

```{r}
addmargins(A = my_table)
```

## Bar graphs

Bar graphs can be helpful for exploring tables

::: columns
::: {.column width="50%"}
```{r}
ggplot(data = assortive_mating, mapping = aes(x=self_male)) +
  geom_bar()
```
:::

::: {.column width="50%"}
```{r}
ggplot(data = assortive_mating, mapping = aes(x=partner_female)) +
  geom_bar()
```
:::
:::

## Mosaic plots

In the case of two categorical variables the mosaic plot is nice.

```{r}
# Here I am showing that I want to see a relationship between the two variables. 
mosaicplot(self_male ~ partner_female, data=assortive_mating)
```

## 2 variables with bar

We can map aes to multiple variables

```{r}
#| code-line-numbers: "1"
ggplot(data = assortive_mating, mapping = aes(x=self_male, fill=partner_female)) +
  geom_bar()
```

## fill

```{r}
#| code-line-numbers: "2"
ggplot(data = assortive_mating, mapping = aes(x=self_male,fill=partner_female)) +
  geom_bar(position = "fill")
```

## dodge

```{r}
#| code-line-numbers: "3"
ggplot(data = assortive_mating, mapping = aes(x=self_male,fill=partner_female)) +
  geom_bar(position = "dodge")
```

## [Problem 4.5](https://openintro-ims.netlify.app/explore-categorical#chp4-exercises)

## Additional Practice exploring Categorical Data

Once you've loaded the openintro and tidyverse libraries there are many dataset to practice with. 

`avandia` is the name of the dataset explore it using the methods above. 

I'll explore it myself in the next video.

