---
title: "CI practice"
author: "Schwab"
format: revealjs
editor: visual
execute: 
  echo: true
---

# Libraries

```{r}
library(openintro)
library(tidyverse)
```

# Problem 1

```{r}
age_at_mar |> head(3)
```

Two questions before doing the CLT

1.  Is the sample large?
2.  Are the data Values independent?

# Make a 95% CI

```{r}

mean(age_at_mar$age, na.rm =TRUE)
sd(age_at_mar$age, na.rm =TRUE)
```

$\bar{X} \sim N(\mu,\frac{\sigma}{\sqrt{n}})$

$\bar{x} \pm z^*SE$

Make a 90% CI

Interpret the CI

# Problem 2

NYC marathon winning times for women

```{r}
set.seed(10252023)

nyc_marathon_women_sample <- nyc_marathon |> 
  filter(division == "Women" & time_hrs <2.58 ) |>
  select(time_hrs) |>
  slice_sample(n=10)
```

```{r}
#| echo: false
nyc_marathon |> 
  filter(division == "Women" ) |>
  select(time_hrs) |>
  ggplot(aes(time_hrs))+
  geom_boxplot()
```

# Find a 99.7% CI

Find a 99.7% CI for the NYC winning woman's time.

Find the mean and SD for `nyc_marathon_women_sample`

Draw Picture

Make interval

Interpret

# We know the pop mean

```{r}
nyc_marathon |> 
  filter(division == "Women" & time_hrs <2.58 ) |>
  summarise(ave = mean(time_hrs))

```

# Other data sets and problems:

1.  `climber_drugs`\

-   What is the average amount of Acetazolamide? Make a 80% CI

2.  `get_it_dunn_run`\

-   CI for average run time.

3.  `heart_transplant`

-   92% CI for survive time.

# `nba_heights`
