---
title: "CLT proportions"
format: revealjs
editor: visual
---

```{r}
#| echo: false
#| include: false


library(tidyverse)
library(openintro)
```

<!-- # Notes on normality -->

<!-- Yesterday we looked at `climber_drugs`, `get_it_dunn_run`, `heart_transplant` and made confidence intervals. -->

<!-- But should we have? -->

<!-- Normality: -->

<!-- -   small sample sizes should have no outliers. We want the data distribution to be normal. -->

<!-- -   large sample sizes should have no *extreme* outliers. It doesn't really matter what the data distribution is. -->

<!-- Let's check those for outliers. -->

<!-- # `climber_drugs` -->

<!-- ```{r} -->
<!-- climber_drugs |> -->
<!--   filter(substance == "Acetazolamide") |> -->
<!--   ggplot(aes(x=concentration -->
<!--   ))+ -->
<!--   geom_boxplot() -->
<!-- ``` -->

<!-- # `get_it_dunn` -->

<!-- ```{r} -->
<!-- get_it_dunn_run |> -->
<!--   ggplot(aes(run_time_minutes))+ -->
<!--   geom_boxplot() -->
<!-- ``` -->

<!-- # `get_it_dunn` part 2 -->

<!-- ```{r} -->
<!-- get_it_dunn_run |> -->
<!--   ggplot(aes(run_time_minutes, fill= sex))+ -->
<!--   geom_boxplot() -->
<!-- ``` -->

<!-- # `get_it_dunn` part 3 -->

<!-- ```{r} -->
<!-- get_it_dunn_run |> -->
<!--   ggplot(aes(run_time_minutes, fill= sex, color=race))+ -->
<!--   geom_boxplot() -->
<!-- ``` -->

<!-- # `heart_transplant` -->

<!-- ```{r} -->
<!-- heart_transplant |> -->
<!--   ggplot(aes(x=survtime))+ -->
<!--   geom_boxplot() -->
<!-- ``` -->

# The Central Limit Theroem 



# CLT Proportions

Same idea as the CLT for a single mean.

We want to create a theoretical sampling distribution entered around $p$ and with a standard deviation of *SE*



# Notation

$$ \hat{p} \sim N(p,SE)$$

With $SE= \sqrt{\frac{p(1-p)}{n}}$

# Rules of thumb

::: {incremental}
-   The operations need to be independent.

-   n should be large

-   Success-Failure Condition

    -   $np >=10$ and

    -   $n(1-p) >= 10$
:::

## Quarantine

We ask 1042 New Yorkers if they are for quarantining peope who have been exposed to ebola. The results are below.

```{r}
count(ebola_survey,quarantine)
```

-   Write the notation for the theoretical sampling distribution.

-   Make a 90, 95 and 99.7% CI for those against and interpret.

## Exclusive relationships

What is the proportion of college students that have had more than 1 `exclusive_relationships`?

```{r}
exclusive_relationship |>
  count(num > 1)

```

-   Check Conditions.

-   Write the notation for the theoretical sampling distribution.

-   Make a 90, 95 and 99.7% CI for those against and interpret.

## Uniforms

A student polls his school to see if students in the school district are for or against the new legislation regarding school uniforms. She surveys 600 students and finds that 480 are against the new legislation.

a.  Compute a 90% confidence interval for the true percent of students who are against the new legislation, and interpret the confidence interval.

## iphones

b.  In a sample of 300 students, 68% said they own an iPod and a smart phone. Compute a 97% confidence interval for the true percent of students who own an iPod and a smartphone.[^1]

[^1]: This and the previous problem were from a text called [Introductory Statistics](https://openstax.org/books/introductory-statistics/pages/8-3-a-population-proportion)

## Consider `male_heights`

If we are interested in the average male height. Should we use:

$$
\hat{p} \sim N(p,\sqrt{\frac{p(1-p)}{n}})
$$

## Exclusive relationships again

If we are considering the average number of exclusive relationships which sampling distribution should we use?

Can you find a 95% CI for the average number of exclusive relationships?

```{r}
exclusive_relationship 
```
