---
title: "Multiple Predictors"
author: "Schwab"
format: revealjs
editor: visual
---

## Libraries

```{r}
library(openintro)
library(tidyverse)
library(infer)
library(broom)
library(MASS)
library(kableExtra)
```

# Linear Regression with cat-egorical data

## Recall the cats data

```{r}
set.seed(12042023)

sample_n(cats, size=7) |>
  head(7)|>
  kable()
```

## Is `sex` a good predictor of `Hwt`?

```{r}
ggplot(data= cats, aes(Sex, Hwt))+
  geom_point()
```

## Math Model

```{r}
lm(Hwt ~ Sex, data = cats) |>
tidy(conf.int = T, conf.level = .95)  |>
  kable(digits = 2)
```

<br>

The mean `Hwt` when `Sex`= F is 9.2.

We expect the mean to rise by 2.12 grams when `Sex=` M

Sex is called an indicator variable. It is binary.

## Try one:

Is smoking a predictor of birth weight? Use `births14`

We'll focus on interpreting the slope and intercept.

## Solution:

```{r}
#| echo: FALSE
lm(weight ~ habit, data= births14)|>
  tidy()
```

## Multiple levels in a variable.

Consider the `loans_full_schema` data from openintro.

```{r}
loans_full_schema |>
  dplyr::select(interest_rate, verified_income)
```

## Graph

```{r}
ggplot(data=loans_full_schema, aes(x= verified_income, y= interest_rate))+
  geom_point()
```

Note: the 3 levels in `verified income`

```{r}
lm(interest_rate ~ verified_income, data = loans_full_schema) |>
tidy()|>
  kable(digits = 2)
```
