---
title: "Digestion Day"
author: "Schwab"
format: revealjs
editor: visual
execute: 
  echo: true
---

# Tons of Vocab:

-   Population

    -   Parameter $\mu$

-   Sample

    -   Statistic $\bar{x}$

-   Variable types (numeric, categorical)

-   Observational Study vs Experiments

# Methods of Sampling

::: incremental
-   Simple Random

-   Cluster

-   Stratified

-   Convience \<- Don't do this one in real life.
:::

# Some R Code

::: incremental
-   `ggplot(data= , aes(x,y)) |> geom_point()`

    -   This makes a point plot

-   `present |> mutate(total = x + y)`

    -   This adds a cloumn

-   `present |> arrange(desc(total))`

    -   This arranges the total colum in descending order.

-   `$` and `?`

    -   Pulls a column out of a df and get the help menu respectively
:::

# Tutorial R things

::: incremental
-   `hsb2 |>   count(schtyp)`

-   `hsb2_public <- hsb2 |>   filter(schtyp == "public")`

-   `if_else()` and `case_when()` functions

-   `group_by()`

-   numeric variables are \<int\> and \<dbl\>

-   categorical variables are \<fct\> and \<chr\>
:::

# Live coding example

Load tidyverse and openintro

Consider the `babies` df

# Don't ruin the surprise

If you click the next slide you will see the code I hope to do.

# Code I hope to do:

```{r}
#| echo: false
  library(tidyverse)
  library(openintro)
```
# babies
```{r}
#| echo: true
  babies_lbw <- babies |>
    mutate(bwt_type= 
             case_when(
              bwt < 3.23 ~ "very low birth weight",
              bwt > 3.23 &  bwt < 5.5 ~ "low birth weight",
              bwt >= 5.5 ~ "healthy birth weight"
    )
    )
  babies_lbw
```
# bargraph?
```{r}
babies_lbw |>
  ggplot(aes(x=smoke)) +
  geom_bar(aes(fill=bwt_type))
```

# count
```{r}
babies_lbw |>
  count(bwt)
```

# Take 2
```{r}

babies_lbw <- babies |>
  mutate(bwt_type= 
           case_when(
            bwt < 3.23*16 ~ "very low birth weight",
            bwt > 3.23*16 &  bwt < 5.5*16 ~ "low birth weight",
            bwt >= 5.5*16 ~ "healthy birth weight"
    )
  )
# Numbers are multiplied by 16 to give ounces from pounds
  
  
```

# new bargraph
```{r}
babies_lbw |>
  ggplot(aes(x=smoke)) +
  geom_bar(aes(fill=bwt_type), position = "fill")
```
