---
title: " R Assignment 6"
subtitle:  "Test for a difference of means"
author: "name"
date: "`r Sys.Date()`"
output: pdf_document
execute: 
  echo: true
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(eval = TRUE)
library(tidyverse)
library(openintro)

<<<<<<< HEAD:MTH_142_files_for_students/RAssignments/R_Assignment_5.rmd

=======
>>>>>>> b5fb0b334a31d3d7690277c744991d20e0ba2a95:old R Assignments/R_Assignment_6_test_diff_of_means.rmd
```

## Assignmnet

<<<<<<< HEAD:MTH_142_files_for_students/RAssignments/R_Assignment_5.rmd
In this final R Assignment you will make a confidence interval for a single mean. We will see if the confidence would have captured the true population parameter. 

## The data

There was an episode of the Planet Money podcast in which people guessed the weight of a cow named Penelope. I'd like to know if a confidence interval around the average of the weight guessed would have actually captured the population parameter, which is Penelope's actual weight of 1355 pounds. Type ?penelope to pull up the documentation to learn more and follow the link to hear the podcast. 

```{r}
# note there is code to make a histogram in the documentation, feel free to use it
=======
In this final R Assignment you will carry out a hypothesis test for a single mean. The instructions are similar to that of the previous two R Assignments. You have gotten better at R so I am telling you a little less. 



## The data

There was an episode of the Planet Money podcast in which people guessed the weight of a cow named Penelope. You should recall this data from an earlier R Assignment. We know the true average I'd like to know if the average of the weights guessed was statistically close to Penelope's actual weight of 1355 pounds. Type ?penelope to learn more and hear the podcast. 

```{r}
smoking |>
  filter(smoke =="yes")
>>>>>>> b5fb0b334a31d3d7690277c744991d20e0ba2a95:old R Assignments/R_Assignment_6_test_diff_of_means.rmd

```


## Do an exploritory analysis. 

At minimum you should make a graph to look at the data, and also find the average guessed weight of the cow. This data is numeric, so you should not make a bar graph, reference RAssignment 1 for the correct graph to check for outliers. Feel free to find other information that might be necessary for conditions.  

```{r}

```

## No outliers

Some of the guesses are very small small/ large, but Planet Money already removed the outliers from this data. So the dots in a boxplot above are not actual outliers. R is calculating outliers assuming this is the original data (if you made a boxplot add the argument coef = 3 to cover them up). So we can proceed with our test.If you have questions about this please post to the discussion and I'll explain further. 


## Check conditions

check the conditions for making a confidence interval. 



## Make the 99% confidence interval. 

Use t.test() to make a confidence interval ( reference the lecture from HW 8). 


```{r}


```

## Questions

### Question 1.

99% of the time we would expect Penelope's actual weight to have fallen in the confidence interval. Did Penelope's weight fall in the range of the confidence interval? Justify by stating both your confidence interval and the population mean. 

### Question 2

Assuming that your work is correct, then either our conditions were not met or we just happened to get a sample that did not contain the population parameter (which happens 1% of the time). If any conditions were not met (besides outliers) speculate as to why they were not met in terms of people guessing. You may want to consider the [anchoring effect](https://en.wikipedia.org/wiki/Anchoring_effect). You may want to take a look at the actual numbers in the data by putting view(penelope) in a code chunk and running it.  


### Question 3

The guessers were not that far off, what percent incorrect were they? To calculate this do:

$$|\frac{\bar{x}-\mu}{\mu}|$$

If you do the above calculation in R abs() will find the absolute value of a number. So even though statistically they were incorrect,  practically speaking they were pretty close. 

